%\paragraph{Practical considerations \todo{Move to perspectives}}
%A component does not execute on top of a flat operating system process but on top of a managed runtime environment, the JVM.
%As a consequence, we devote resources to the component but also to the middleware.
%Due to the presence of the garbage collector, we do not know the exact amount of memory new \textit{JVM nodes} should be assigned with.
%In fact, we can only state that the reserved memory must equal the component reservation plus the middleware requirement plus any amount that allows an efficient behavior %of the garbage collector.
%Chances are that any prediction will overestimate the real consumption, leading to the problem of wasting resources due to internal fragmentation.

\section{Conclusion} \label{sec:conclusions}

%\enlargethispage{0.3cm}

This paper presents Squirrel, a framework that provides resource management capabilities to dynamic component-based frameworks.
Squirrel proposes choosing component-to-system mappings at deployment time for better resource management.
This strategy is performed automatically by checking the resource availability and transforming the application's structure to run the application on resource-aware containers.
Containers describe how to map components to system abstractions
%in ways that reduce the cost of resource management.
allowing for different trade-offs in resource management.

We present an implementation of Squirrel that manages CPU, I/O and memory, and provide performance analyses and a comparison of different design decisions.
The experiments
%to assess our implementation's behavior 
show that choosing the right component-to-system mappings at deployment-time reduces CPU overhead and/or memory use.
They also highlight that optimizing mappings is essential to reducing isolation and communication overhead to acceptable levels.
%and make it affordable.

In the future, we envision augmenting the Squirrel framework with automatic and dynamic reasoning capabilities in order to automatically place and migrate components with resource contracts over a distributed architecture.
%This paper also opens interesting perspectives by applying the presented concepts in different application domains with strong security and safety concerns.
Applying the concepts in this paper to other application domains with strong safety and security concerns is of interest.
For instance, security concerns often leverage resource isolation as a means to avoid sharing confidential information with untrusted components within the same container.
%\hl{WHAT?: Indeed, security concerns are often correlated with resource isolation in managing that some confidential information cannot be shared to untrusted components within the same container. }
The possibility of applying a pattern-based approach to increase an application's security without compromising its efficiency looks promising~\cite{DBLP:conf/kbse/MorinMFTBJ10}.



%Energy efficient dynamic adaptation, to switch off parts of the electronics depending on the resources usage